---
title: "Modeling of Development Data"
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE, 
                      message = FALSE)
```

```{r}
library(RODBC)
library(boot)
library(tidyverse)
library(tidymodels)
library(lmtest)
library(randomForest)
library(xgboost)

# Figure options
theme_set(theme_bw())
theme_update(panel.grid = element_blank())
```

```{r, data}

# Using frozen data from 09/13/2024 data transfer
motrpac <- odbcConnect(dsn = "motrpac_XFER_RP",
                       uid = Sys.getenv("motrpac_uid"),
                       pwd = Sys.getenv("motrpac_pwd")) 

# Bring in the development data combined in SAS
dev_data <- motrpac %>%
  sqlQuery('SELECT * FROM SASRampProtocol')

#close ODBC connection
odbcCloseAll()
```

## Scatterplots to pick functional forms

```{r}

ggplot(data = dev_data, aes(x = grip, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Grip Strength, kg") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = peaktorqavg_iske, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Isometric Knee Extension, Newton Meters") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = BSA, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab(bquote("Body Surface Area, " ~m^2)) + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = age_psca, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Age, years") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = waist, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Waist Circumference, cm") + 
  ylab("Watts at Termination")
```


\newpage

```{r, impute}
dev_data_reduced <- dev_data %>%
  select(-c("chosenprotocol", "inrange", "duration"))

# impute missing data in peaktorqavg_iske and grip strength variables using
# k-nearest neighbors
imputer=recipe(wattend_cpet~., data=dev_data_reduced)%>%
  step_rm(pid)%>%
  step_string2factor(class)%>%
  step_dummy(class)%>%
  step_impute_knn(peaktorqavg_iske,grip)%>%
  prep(training=dev_data_reduced)

dev_data_i=juice(imputer)

```


## Linear Regression

### Data for Table 4 - Coefficients from the final linear regression model to predict watts at conclusion of the cardiopulmonary exercise test.

```{r}
lm1=lm(wattend_cpet ~ BSA+ peaktorqavg_iske + sex2 +age_psca + class_HARE + class_SED + waist + grip + hr, data=dev_data_i)
summary(lm1)


plot(dev_data$wattend_cpet~(predict(lm1, dev_data_i)), 
     main="Predicted vs Observed Watts with Linear Regression", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination"
)

to_calc_rsq <- dev_data_i %>%
  mutate(predicted = predict(lm1, dev_data_i))
  
rsq(to_calc_rsq, wattend_cpet, predicted)

lm_onlyAge=lm(wattend_cpet ~ age_psca , data=dev_data_i)

lrtest(lm1, lm_onlyAge)
```


\newpage

## Random Forest 

```{r}
ramp_rf <- dev_data_i %>% 
  mutate_if(is.character, as.factor)

rf <- randomForest(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + class_HARE+class_SED+waist +grip +hr,
                   data = ramp_rf,  
                   importance=TRUE)

plot(dev_data_i$wattend_cpet~ (predict(rf, ramp_rf)), 
     main="Predicted vs Observed Watts with Random Forest", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination"
)

```


\newpage

## Gradient Boosting 

```{r}
ramp_xg <- data.matrix(select(ramp_rf, - wattend_cpet))

boost = xgboost(data = ramp_xg, 
                label = ramp_rf$wattend_cpet, 
                objective = "reg:squarederror", 
                nrounds=10)

plot(dev_data_i$wattend_cpet ~ (predict(boost, ramp_xg)), 
     main="Predicted vs Observed Watts with Extreme Gradient Boosting", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination")

```


\newpage

## Cross-Validation

```{r}

set.seed(865)
folds <- sample(rep(1:10, 40) )
dev_data_i$fold <- folds
ramp_rf$fold <- folds


MSElm=rep(NA, 10)
MSErf=rep(NA, 10)
MSExg=rep(NA, 10)

MSElm_onlyAge = rep(NA, 10)

RSQlm=rep(NA, 10)
RSQrf=rep(NA, 10)
RSQxg=rep(NA, 10)

RSQlm_onlyAge = rep(NA, 10)


for (i in 1:10){
  ts=subset(dev_data_i,fold!=i)
  vs=subset(dev_data_i, fold==i)
  ts2=subset(ramp_rf,fold!=i)
  vs2=subset(ramp_rf, fold==i)
  tsxg=data.matrix(select(ts2, - wattend_cpet))
  vsxg=data.matrix(select(vs2, - wattend_cpet))

  # Calculate MSE
  fitlm=lm(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + class_HARE+class_SED +waist+ grip+ hr, data=ts)
  MSElm[i]=mean((vs$wattend_cpet-predict(fitlm, vs))^2)
  
  fitlm_onlyAge=lm(wattend_cpet ~ age_psca , data=ts)
  MSElm_onlyAge[i]=mean((vs$wattend_cpet-predict(fitlm_onlyAge, vs))^2)
  
 fitrf <- randomForest(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + class_HARE+class_SED+waist +grip +hr, data=ts2,  importance=TRUE,)
   MSErf[i]=mean((vs2$wattend_cpet-predict(fitrf, vs2))^2)
  
     fitxg = xgboost(data = tsxg, label = ts2$wattend_cpet, objective = "reg:squarederror", nrounds=10, verbose=0)
     vsxgdf <- as.data.frame(vsxg)
   MSExg[i]=mean((vs2$wattend_cpet-predict(fitxg, vsxg))^2)

   # Calculate rsq
 to_calc_rsqlm <- vs %>%
    mutate(predlm = predict(fitlm, vs),
           predlm_onlyAge = predict(fitlm_onlyAge, vs))

 to_calc_rsqrf <- vs2 %>%          
    mutate(predrf = predict(fitrf, vs2))
 
 to_calc_rsqxg <- vs2
 to_calc_rsqxg$predxg = predict(fitxg, vsxg)
  
 RSQlm[i] <- rsq(to_calc_rsqlm, wattend_cpet, predlm)$.estimate
 RSQlm_onlyAge[i] <- rsq(to_calc_rsqlm, wattend_cpet, predlm_onlyAge)$.estimate

 RSQrf[i] <- rsq(to_calc_rsqrf, wattend_cpet, predrf)$.estimate
 RSQxg[i] <- rsq(to_calc_rsqxg, wattend_cpet, predxg)$.estimate

 
}


```

### Data for Table 3 - Evaluations of model performances in development data. 10-fold cross-validation used to estimate mean squared error and R-squared.

The MSE for the full linear model is `r round(mean(MSElm),digits = 2)`.

The MSE for the age only linear model is `r round(mean(MSElm_onlyAge),digits = 2)`

The MSE for the random forest model is `r round(mean(MSErf),digits = 2)`

The MSE for the gradient boosted model is `r round(mean (MSExg),digits = 2)`


The R-squared for the full linear model is `r round(mean(RSQlm),digits = 3)`

The R-squared for the age only model is `r round(mean(RSQlm_onlyAge),digits = 3)`

The R-squared for the random forest model is `r round(mean(RSQrf),digits = 3)`

The R-squared for the gradient boosted model is `r round(mean(RSQxg),digits = 3)`