---
title: "Modeling of Development Data"
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE, 
                      message = FALSE)
```

```{r}
library(RODBC)
library(boot)
library(tidyverse)
library(tidymodels)
library(lmtest)
library(randomForest)
library(xgboost)
library(ggplot2)
library(here)

# Figure options
theme_set(theme_bw())
theme_update(panel.grid = element_blank())
```

```{r, data}

# Using frozen data from 09/13/2024 data transfer
motrpac <- odbcConnect(dsn = "motrpac_XFER_RP",
                       uid = Sys.getenv("motrpac_uid"),
                       pwd = Sys.getenv("motrpac_pwd")) 

# Bring in the development data combined in SAS
dev_data <- motrpac %>%
  sqlQuery('SELECT * FROM SASRampProtocol')

# We also want BMI now too
bl_bmi_data <- motrpac %>%
  sqlQuery('SELECT pid, visitcode, calcbmi_hwwt
           FROM v_HWWT') %>%
  filter(visitcode == "ADU_SCP" & pid %in% dev_data$pid) %>%
  select(-visitcode)

#close ODBC connection
odbcCloseAll()

dev_data_w_BMI <- left_join(dev_data, bl_bmi_data, by = "pid")

```

## Scatterplots to pick functional forms

```{r}

ggplot(data = dev_data, aes(x = grip, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Grip Strength, kg") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = peaktorqavg_iske, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Isometric Knee Extension, Newton Meters") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = BSA, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab(bquote("Body Surface Area, " ~m^2)) + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = age_psca, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Age, years") + 
  ylab("Watts at Termination")

ggplot(data = dev_data, aes(x = waist, y = wattend_cpet)) +
  geom_point() + 
  geom_smooth(color = 'black') +
  xlab("Waist Circumference, cm") + 
  ylab("Watts at Termination")
```


\newpage

```{r, impute}
dev_data_reduced <- dev_data %>%
  select(-c("chosenprotocol", "inrange", "duration"))

# impute missing data in peaktorqavg_iske and grip strength variables using
# k-nearest neighbors
imputer=recipe(wattend_cpet~., data=dev_data_reduced)%>%
  step_rm(pid)%>%
  step_string2factor(class)%>%
  step_dummy(class)%>%
  step_impute_knn(peaktorqavg_iske,grip)%>%
  prep(training=dev_data_reduced)

dev_data_i=juice(imputer)

```


## Linear Regression

### Data for Table 4 - Coefficients from the final linear regression model to predict watts at conclusion of the cardiopulmonary exercise test.

```{r}
lm1=lm(wattend_cpet ~ BSA+ peaktorqavg_iske + sex2 +age_psca + class_HARE + class_SED + waist + grip + hr, data=dev_data_i)
summary(lm1)

to_calc_rsq <- dev_data_i %>%
  mutate(predicted = predict(lm1, dev_data_i))
  
rsq(to_calc_rsq, wattend_cpet, predicted)

# No peaktorque or grip strength model
lm2=lm(wattend_cpet ~ BSA + sex2 +age_psca + class_HARE + class_SED + waist + hr, data=dev_data_i)

# Age only model
lm_onlyAge=lm(wattend_cpet ~ age_psca , data=dev_data_i)

lrtest(lm1, lm_onlyAge)

# Age, BMI, and sex model, restingHR & physical activity measurement
# dev_data_i_w_BMI <- left_join(dev_data_i, 
#                               select(dev_data_w_BMI, -peaktorqavg_iske, -grip)) %>%
#   select(age_psca, sex2, BSA, wattend_cpet, hr, class_HARE, class_SED)
# 
# lm_reduced_vars=lm(wattend_cpet ~ age_psca + sex2 + calcbmi_hwwt + hr+ class_HARE + class_SED, data=dev_data_i_w_BMI)


```

\newpage

### Linear Regression Observed vs Predicted plot

```{r}
for_plot <- dev_data_i %>%
  mutate(predicted = predict(lm1, dev_data_i))

for_plot %>%
  ggplot(aes(x = predicted, y = wattend_cpet)) +
  geom_point(size = .75) +
  geom_abline(slope = 1, intercept =0 ) +
  ylab("Observed Watts at Termination") + 
  xlab("Predicted Watts at Termination") +
  ylim(c(50,475)) +
  xlim(c(50,475))
  

ggsave(file.path(here(), "reports", "Figures", "Supplemental_Fig_3.tif"),
     device = "tiff", dpi = 600, width = 5.75, height = 4, units = "in")
```

\newpage

### Data for Supplemental - Coefficients from the linear regression model without peak torque and grip strength to predict watts at conclusion of the cardiopulmonary exercise test.

```{r}
lm2=lm(wattend_cpet ~ BSA + sex2 +age_psca + class_HARE + class_SED + waist + hr, data=dev_data_i)
summary(lm2)


plot(dev_data$wattend_cpet~(predict(lm2, dev_data_i)), 
     main="Predicted vs Observed Watts with Linear Regression", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination"
)

```

\newpage

## Random Forest 

```{r}
ramp_rf <- dev_data_i %>% 
  mutate_if(is.character, as.factor)

rf <- randomForest(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + class_HARE+class_SED+waist +grip +hr,
                   data = ramp_rf,  
                   importance=TRUE)

plot(dev_data_i$wattend_cpet~ (predict(rf, ramp_rf)), 
     main="Predicted vs Observed Watts with Random Forest", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination"
)

```


\newpage

## Gradient Boosting 

```{r}
ramp_xg <- data.matrix(select(ramp_rf, - wattend_cpet))

boost = xgboost(data = ramp_xg, 
                label = ramp_rf$wattend_cpet, 
                objective = "reg:squarederror", 
                nrounds=10)

plot(dev_data_i$wattend_cpet ~ (predict(boost, ramp_xg)), 
     main="Predicted vs Observed Watts with Extreme Gradient Boosting", 
     xlab="Predicted Watts at Termination", 
     ylab="Observed Watts at Termination")

```


\newpage

## Cross-Validation

```{r}

set.seed(865)
folds <- sample(rep(1:10, 40) )
dev_data_i$fold <- folds
# dev_data_i_w_BMI$fold <- folds
ramp_rf$fold <- folds


MSElm=rep(NA, 10)
MSErf=rep(NA, 10)
MSExg=rep(NA, 10)

MSElm2=rep(NA, 10)
MSElm_onlyAge = rep(NA, 10)
MSElm_reducedVars = rep(NA, 10)

RSQlm=rep(NA, 10)
RSQrf=rep(NA, 10)
RSQxg=rep(NA, 10)

RSQlm2=rep(NA, 10)
RSQlm_onlyAge = rep(NA, 10)
RSQlm_reducedVars = rep(NA, 10)

for (i in 1:10){
  # Subsets for Linear modeling
  ts=subset(dev_data_i,fold!=i)
  vs=subset(dev_data_i, fold==i)
  
  # ts_reduced=subset(dev_data_i_w_BMI,fold!=i)
  # vs_reduced=subset(dev_data_i_w_BMI, fold==i)
  
  
  # Subsets for Random forest
  ts2=subset(ramp_rf,fold!=i)
  vs2=subset(ramp_rf, fold==i)
  
  # Subsets for Gradient boosting
  tsxg=data.matrix(select(ts2, - wattend_cpet))
  vsxg=data.matrix(select(vs2, - wattend_cpet))

  # Calculate MSE
  # All predictors LM
  fitlm=lm(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + class_HARE+
             class_SED + waist + grip + hr, data=ts)
  MSElm[i]=mean((vs$wattend_cpet-predict(fitlm, vs))^2)
  
  # Removing peaktorq and grip strength
  fitlm2=lm(wattend_cpet ~ BSA +sex2 +age_psca + class_HARE+
             class_SED + waist + hr, data=ts)
  MSElm2[i]=mean((vs$wattend_cpet-predict(fitlm2, vs))^2)
  
  # Only age LM
  fitlm_onlyAge=lm(wattend_cpet ~ age_psca , data=ts)
  MSElm_onlyAge[i]=mean((vs$wattend_cpet-predict(fitlm_onlyAge, vs))^2)
  
  # Reduced predictors (with BMI)
  # fitlm_reducedVars=lm(wattend_cpet ~ age_psca + sex2 + calcbmi_hwwt + hr+ class_HARE + class_SED, 
  #                      data=ts_reduced)
  # MSElm_reducedVars[i] = mean((vs$wattend_cpet - 
  #                                predict(fitlm_reducedVars, vs_reduced))^2)
  
  # Random Forest
  fitrf <- randomForest(wattend_cpet ~ BSA+ peaktorqavg_iske +sex2 +age_psca + 
                      class_HARE+class_SED+waist +grip +hr, 
                      data=ts2,  importance=TRUE,)
  MSErf[i]=mean((vs2$wattend_cpet-predict(fitrf, vs2))^2)
  
  #Gradient boosting
  fitxg = xgboost(data = tsxg, label = ts2$wattend_cpet, 
                  objective = "reg:squarederror", nrounds=10, verbose=0)
  vsxgdf <- as.data.frame(vsxg)
  MSExg[i]=mean((vs2$wattend_cpet-predict(fitxg, vsxg))^2)

  # Calculate r-squared
  # Linear models
  to_calc_rsqlm <- vs %>%
    mutate(predlm = predict(fitlm, vs),
           predlm2 = predict(fitlm2, vs),
           predlm_onlyAge = predict(fitlm_onlyAge, vs))

  # to_calc_rsqlm_reduced <- vs_reduced %>%
  #   mutate(predlm_reducedVars = predict(fitlm_reducedVars, vs_reduced))

  # Random forest
  to_calc_rsqrf <- vs2 %>%          
    mutate(predrf = predict(fitrf, vs2))
 
  # Gradient boosting
  to_calc_rsqxg <- vs2
  to_calc_rsqxg$predxg = predict(fitxg, vsxg)
  
  
  RSQlm[i] <- rsq(to_calc_rsqlm, wattend_cpet, predlm)$.estimate
  RSQlm2[i] <- rsq(to_calc_rsqlm, wattend_cpet, predlm2)$.estimate
  RSQlm_onlyAge[i] <- rsq(to_calc_rsqlm, wattend_cpet, predlm_onlyAge)$.estimate
  # RSQlm_reducedVars[i] <- rsq(to_calc_rsqlm_reduced, wattend_cpet, predlm_reducedVars)$.estimate

  
  RSQrf[i] <- rsq(to_calc_rsqrf, wattend_cpet, predrf)$.estimate
  RSQxg[i] <- rsq(to_calc_rsqxg, wattend_cpet, predxg)$.estimate

 
}


```

### Data for Table 3 - Evaluations of model performances in development data. 10-fold cross-validation used to estimate mean squared error and R-squared.

The MSE for the full linear model is `r round(mean(MSElm),digits = 2)`.

The MSE for the linear model without peak torque or grip strength is `r round(mean(MSElm2),digits = 2)`.


The MSE for the age only linear model is `r round(mean(MSElm_onlyAge),digits = 2)`


The MSE for the random forest model is `r round(mean(MSErf),digits = 2)`

The MSE for the gradient boosted model is `r round(mean (MSExg),digits = 2)`


The R-squared for the full linear model is `r round(mean(RSQlm),digits = 3)`

The R-squared for the age only model is `r round(mean(RSQlm_onlyAge),digits = 3)`

The R-squared for the linear model without peak torque or grip strength is `r round(mean(RSQlm2),digits = 3)`.


The R-squared for the random forest model is `r round(mean(RSQrf),digits = 3)`

The R-squared for the gradient boosted model is `r round(mean(RSQxg),digits = 3)`