---
title: "Summary of Development Data"
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE, 
                      message = FALSE)
library(RODBC)
library(here)
library(tidyverse)
library(gtsummary)
library(labelled)
library(flextable)

source(file.path(here(), "key.R"))
lapply(list.files(file.path(here(), "R"), full.names = TRUE), source)

# Figure options
theme_set(theme_bw())
theme_update(panel.grid = element_blank())
```


```{r, data}

# Using frozen data from 09/13/2024 data transfer
motrpac <- odbcConnect(dsn = "motrpac_XFER_RP",
                       uid = Sys.getenv("motrpac_uid"),
                       pwd = Sys.getenv("motrpac_pwd")) 

# Bring in the development data combined in SAS
dev_data <- motrpac %>%
  sqlQuery('SELECT * FROM SASRampProtocol')

# We also want BMI now too
bl_bmi_data <- motrpac %>%
  sqlQuery('SELECT pid, visitcode, calcbmi_hwwt
           FROM v_HWWT') %>%
  filter(visitcode == "ADU_SCP" & pid %in% dev_data$pid) %>%
  select(-visitcode)
```

 
```{r, prep_table1}

pids_w_site <- motrpac %>%
  sqlQuery('SELECT * FROM v_participantsadults_live') %>%
  mutate(site_orig = floor(siteID/10)*10,
         site_orig = factor(site_orig, levels = siteID_key,
                                       labels = c("A", "B", "C", "D", "E",
                                                  "F", "G", "H", "I", "J", 
                                                  "K", "L")
                         )
  ) %>%
  select(pid, site_orig)

#close ODBC connection
odbcCloseAll()

table1 <- left_join(dev_data, pids_w_site, by = "pid") %>%
  left_join(bl_bmi_data, by = "pid") %>%
  mutate(sex2 = factor(sex2, 
                       levels=c(1,0),
                       labels=c("Female", "Male")),
         chosenprotocol = factor(chosenprotocol,
                                 levels = c(2510, 5015, 5020,
                                            7520, 7525, 10025),
                                 labels=c("25+10","50+15","50+20",
                                          "75+20","75+25","100+25"))
  ) %>%
  set_variable_labels(
    sex2="Sex",
    age_psca="Age, years",
    chosenprotocol = "Chosen protocol",
    BSA = "Body surface area, m^2",
    calcbmi_hwwt = "BMI, kg/m^2",
    waist="Waist circumference, cm",
    peaktorqavg_iske="Average peak torque, Nm",
    grip="Grip strength, kg",
    hr="Resting heart rate, bpm",
    duration="Test duration, minutes",
    inrange="Duration within 8-12 minutes",
    site_orig = "Site",
    class="Study group"
)

# Test for normality to report median (IQR) if variable is not normal
continuous_vars <- c("age_psca", "BSA", "calcbmi_hwwt", "waist", "peaktorqavg_iske", 
                     "grip", "hr", "duration")

test_normality_p_val <- map_dfr(continuous_vars,
                                ~ tibble(Variable = .x,
                                         p.value = shapiro.test(na.omit(table1[[.x]]))$p.value)) %>%
  mutate(non_normal = ifelse(p.value < 0.05, 1, 0)) 

```

```{r, table1}

# Print table 1
table1 %>% 
  select(age_psca, sex2, chosenprotocol, BSA, calcbmi_hwwt, 
         waist, peaktorqavg_iske, grip, hr,
         duration, inrange, site_orig, class) %>%
  tbl_summary(by = class,
              statistic = list(c("BSA", "hr") ~ "{mean} ({sd})",
                               c("age_psca", "calcbmi_hwwt", "waist", "peaktorqavg_iske",
                                 "grip", "duration")  ~ "{median}\n({p25}, {p75})"),
              digits = list("BSA" ~ c(1,1),
                            "hr" ~ c(0,0),
                            c("calcbmi_hwwt", "waist", "peaktorqavg_iske", "duration") ~ c(1,1,1),
                            c("age_psca", "grip") ~ c(0,0,0))) %>%
    add_overall()  %>%
  gtsummary_polish() %>%
  flextable_autofit(width_max = 7, 
                    prop_used_col_1 = .3) %>%
  add_header_lines("Table 2 - Descriptive statistics of development data by study group.*") %>%
  add_footer_lines("*HAEE=highly active endurance exercise; HARE=highly active resistance exercise; SED=sedentary.")

```


```{r, plot_duration}

protocol_colors <- c(
  "50+20" = "#996633",
  "75+20" = "#00B050",
  "50+15" = "#DC50DC",
  "75+25" = "#CC3232",
  "100+25" = "#002060",
  "25+10" = "#92D050",
  "15+8" = "#9DE0F9"
)


table1 %>%
  ggplot(aes(x = duration, y = wattend_cpet, group = chosenprotocol, color = chosenprotocol)) +
  geom_point() +
  scale_fill_manual(values = protocol_colors) + 
  scale_color_manual(values = protocol_colors) + 
  annotate("rect", 
        xmin = 8, xmax = 12, ymin = -Inf, ymax = Inf, 
        alpha=0.2, fill="gray50")  

```

\newpage

```{r, by_women}
women <- filter(table1, sex2 == "Female")

women %>% 
  select(age_psca, sex2, chosenprotocol, BSA, calcbmi_hwwt, 
         waist, peaktorqavg_iske, grip, hr,
         duration, inrange, site_orig, class) %>%
  tbl_summary(by = class,
              statistic = list(c("BSA", "hr") ~ "{mean} ({sd})",
                               c("age_psca", "calcbmi_hwwt", "waist", "peaktorqavg_iske",
                                 "grip", "duration")  ~ "{median}\n({p25}, {p75})"),
              digits = list("BSA" ~ c(1,1),
                            "hr" ~ c(0,0),
                            c("calcbmi_hwwt", "waist", "peaktorqavg_iske", "duration") ~ c(1,1,1),
                            c("age_psca", "grip") ~ c(0,0,0))) %>%
    add_overall()  %>%
  gtsummary_polish() %>%
  flextable_autofit(width_max = 7, 
                    prop_used_col_1 = .3) %>%
  add_header_lines("Supplemental Digital Content 2 - Descriptive statistics of development data, women only.*") %>%
  add_footer_lines("*HAEE=highly active endurance exercise; HARE=highly active resistance exercise; SED=sedentary.")

```


\newpage


```{r, by_men}
men <- filter(table1, sex2 == "Male")

men %>% 
  select(age_psca, sex2, chosenprotocol, BSA, calcbmi_hwwt, 
         waist, peaktorqavg_iske, grip, hr,
         duration, inrange, site_orig, class) %>%
  tbl_summary(by = class,
              statistic = list(c("BSA", "hr") ~ "{mean} ({sd})",
                               c("age_psca", "calcbmi_hwwt", "waist", "peaktorqavg_iske",
                                 "grip", "duration")  ~ "{median}\n({p25}, {p75})"),
              digits = list("BSA" ~ c(1,1),
                            "hr" ~ c(0,0),
                            c("calcbmi_hwwt", "waist", "peaktorqavg_iske", "duration") ~ c(1,1,1),
                            c("age_psca", "grip") ~ c(0,0,0))) %>%
    add_overall()  %>%
  gtsummary_polish() %>%
  flextable_autofit(width_max = 7, 
                    prop_used_col_1 = .3) %>%
  add_header_lines("Supplemental Digital Content 3 - Descriptive statistics of development data, men only.*") %>%
  add_footer_lines("*HAEE=highly active endurance exercise; HARE=highly active resistance exercise; SED=sedentary.")

```

\newpage

## Number in training dataset with watts at termination between 125 to 140
```{r, n_watts_125_140}

out_of_range_within_watts <- dev_data %>%
  filter(wattend_cpet > 125 & wattend_cpet < 140 & inrange == 0 & class == "SED") %>%
  summarize(n = n())

out_of_range <- dev_data %>%
  filter(inrange == 0 & class == "SED") %>%
  summarize(n = n())
```

`r out_of_range_within_watts$n` SED participants had watts at termination between 125 and 140 and were out of range.

`r out_of_range$n` SED participants were out of range.

Of those out of range, `r (out_of_range_within_watts/out_of_range)*100`% were due to having watts at termination between 125 and 140. 


\newpage

## Mirrored Histogram

```{r, hist}

df_plot_ps <- dev_data %>%
  tidyr::pivot_wider(names_from = inrange, 
                     values_from = wattend_cpet, 
                     names_prefix = "ct_p")


ggplot(df_plot_ps) +
  geom_histogram(bins = 50,   alpha = 0.5,
                 aes(x = ct_p1), 
                 fill = "cornflower blue") +
  geom_histogram(bins = 50, alpha = 0.5,
                 aes(x = ct_p0, 
                     y = - stat(count)),
                 fill = "orange") + 
  scale_y_continuous("Count", label = abs) + 
  scale_x_continuous("Watts at Termination") +
  annotate(geom ="label", x= 300, y=15, label="In Range", color="cornflower blue") + 
  annotate(geom ="label", x= 300, y=-15, label="Out of Range", color="orange") + 
  facet_wrap(~class, ncol=3)

ggsave(file.path(here(), "/reports/Figures/Supplemental_Fig_2.tif"),
      device = "tiff", dpi = 600, width = 5.75, height = 3.75, units = "in")
```






